---
title: "Homework 2 Report"
output: html_document
---


## Name

Iris Quan

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
library(tidyverse)

dataset <- read_csv('maacs.csv.gz')
dataset
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}

# To quickly check if our statement of older children having higher FEV1 values than younger children is true for the dataset, we can find the correlation coefficient between the age variable and the fev1 variable.  If the correlation coefficient is greater than 0, it signifies a positive linear association between age and fev1, suggesting that our statement is true, so we return true. The correlation coefficient being less than or equal to 0 suggests otherwise, so in that case we return false.

# function to check if statement is true
q1 <- function(data) {
  # dropping rows with na in age or fev1 column
  data_nona <- data %>% drop_na(age, fev1)
  # printing correlation coefficient
  print(cor(data_nona$age, data_nona$fev1))
  # true if positive corr coeff, false otherwise
  return(cor(data_nona$age, data_nona$fev1) > 0)
}

q1(dataset)

# The function returned true for assessing the validity of the statement "FEV1 values in children are higher in older children relative to younger children" for our dataset.
```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}
# fitting linear model
lmp2 <- lm(fev1 ~ age, data = dataset)
summary(lmp2)
```


Write your data analysis statement interpreting the regression model here:

```{r}
# The coefficients of our linear regression model fitting fev1 on age returned -0.001199 for the intercept and 0.171162 for the slope. The intercept coefficient of -0.001199 liters is the baseline 1 second forced expiratory volume at age = 0.
# Statement: For every 1 year increase in a child's age, there is an estimated increase of 0.171162 liters in the child's forced expiratory volume in 1 second. 
```


## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results.

NOTE: At least one supporting premise should use a plot.


Write the three supporting premise statements here:

1. Supporting Premise 1: boxplots for rounded fev1 vs ages shows increasing median and quartile ages as fev1 increases

2. Supporting Premise 2: high positive correlation coefficient between age and fev1

3. Supporting Premise 3: scatterplot of age vs fev1 shows positive linear association


## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Supporting Premise 1: boxplots for rounded fev1 vs ages shows increasing median and quartile ages as fev1 increases

supp_premise1 <- function(data) {
  # rounding fev1 to be pseudo categorical, plotting
  data$fev1_round <- round(data$fev1)
  boxplot(age ~ fev1_round, data = data,
          xlab = 'Forced Expiratory Volume in 1 second (liters)', 
          ylab = 'Age (years)', main = 'Age vs. FEV1 Boxplot')
  
  # hypothetical age and fev1
  hypo_age = runif(100, 0, 20)
  hypo_fev1 = hypo_age / 6 + 1 + rnorm(100, 0, 0.5)
  # floor and ceiling on fev1
  hypo_fev1[hypo_fev1 < 1] = 1
  hypo_fev1[hypo_fev1 > 5] = 5
  # rounding hypothetical fev1 to be pseudo categorical, plotting
  hypo_fev1_round <- round(hypo_fev1)
  boxplot(hypo_age ~ hypo_fev1_round, 
          xlab = 'Hypothetical Forced Expiratory Volume in 1 second (liters)', 
          ylab = 'Hypothetical Age (years)', 
          main = 'Hypothetical Age vs. FEV1 Boxplot Assuming Statement True')
}

supp_premise1(dataset)
```



```{r}

## Supporting Premise 2: high (above 0.7) positive correlation coefficient between age and fev1

supp_premise2 <- function(data) {
  # dropping rows with na in age or fev1 column
  data_nona <- data %>% drop_na(age, fev1)
  # printing correlation coefficient
  print(cor(data_nona$age, data_nona$fev1))
  # true if positive corr coeff, false otherwise
  return(cor(data_nona$age, data_nona$fev1) > 0.7)
}

supp_premise2(dataset)
```


```{r}
## Supporting Premise 3: scatterplot of age vs fev1 shows positive linear association

supp_premise3 <- function(data) {
  # plotting scatterplot
  plot(data$age, data$fev1)

  # hypothetical age and fev1
  hypo_age = runif(100, 0, 20)
  hypo_fev1 = hypo_age / 6 + 1 + rnorm(100, 0, 0.5)
  # floor and ceiling on fev1
  hypo_fev1[hypo_fev1 < 1] = 1
  hypo_fev1[hypo_fev1 > 5] = 5
  # plotting hypothetical scatterplot
  plot(hypo_age, hypo_fev1, 
       xlab = 'Age (years)',
       ylab = 'FEV1 (liters)',
       main = 'Hypothetical Age vs. FEV1 Scatterplot Assuming Statement True')
}

supp_premise3(dataset)
```

Execute each of your function and show that the produce the expected output.




## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 

```{r}
## Alternative statement: Children's ages have no impact on their FEV1 values; Older children's FEV1 values are neither higher nor lower than the FEV1 values of younger children.
```


Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.


